---
title: "Appendix"
author:
  - name: "Eric Marcon"
  - name: "Florence Puech"
abstract: >
  Code to reproduce the figures of the main text.
date: "`r format(Sys.time(), '%d %B %Y')`"
url: https://ericmarcon.github.io/JSE_appendix/
github-repo: ericmarcon/JSE_appendix
lang: en-US
bibliography: references.bib
biblio-style: chicago
pdftoc: false
toc-depth: 3
urlcolor: blue
always_allow_html: yes
csquotes: true
output:
  rmdformats::downcute:
    use_bookdown: yes
    lightbox: yes
  bookdown::pdf_book:
    template: latex/template.tex
    citation_package: natbib
    latex_engine: xelatex
    keep_tex: yes
---

```{r DoNotModify, include=FALSE}
### Utilities. Do not modify.
# Installation of packages if necessary
InstallPackages <- function(Packages) {
  InstallPackage <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {
      install.packages(Package, repos="https://cran.rstudio.com/")
    }
  }
  invisible(sapply(Packages, InstallPackage))
}

# Basic packages
InstallPackages(c("bookdown", "formatR", "kableExtra", "ragg"))

# kableExtra must be loaded 
if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "docx") {
  # Word output (https://stackoverflow.com/questions/35144130/in-knitr-how-can-i-test-for-if-the-output-will-be-pdf-or-word)
  # Do not use autoformat (https://github.com/haozhu233/kableExtra/issues/308)
  options(kableExtra.auto_format = FALSE)
}
library("kableExtra")

# Chunk font size hook: allows size='small' or any valid Latex font size in chunk options
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r Options, include=FALSE}
### Customized options for this document
# Add necessary packages here
Packages <- c("dplyr", "ggplot2", "stringr", "sf", "dbmss", "magick")
# Install them
InstallPackages(Packages)

# Crop images (requires package magick 
# and pdfcrop installed by tinytex::tlmgr_install("pdfcrop"))
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
# knitr options
knitr::opts_chunk$set(
  cache =   TRUE,     # Cache chunk results
  crop =    TRUE,     # Crop fig margins in pdf  
  include = TRUE,     # Show/Hide chunks
  echo =    TRUE,     # Show/Hide code
  warning = FALSE,    # Show/Hide warnings
  message = FALSE,    # Show/Hide messages
  # Figure alignment and size
  fig.align = 'center', out.width = '80%',
  # Graphic devices (ragg_png is better than standard png)
  dev = c("ragg_png", "pdf"),
  # Code chunk format
  tidy = TRUE, tidy.opts = list(blank=FALSE, width.cutoff=50),
  size = "scriptsize", knitr.graphics.auto_pdf = TRUE
  )
options(width = 50)

# ggplot style
library("ggplot2")
theme_set(theme_bw())
theme_update(panel.background=element_rect(fill="transparent", colour=NA),
             plot.background=element_rect(fill="transparent", colour=NA))
knitr::opts_chunk$set(dev.args=list(bg="transparent"))

# Random seed
set.seed(973)
```


# Theoretical example

This section allows reproducing the theoretical example presented in the main text.
Analyses rely on the _dbmss_ [@Marcon2014] package for R [@R].


## Dataset simulation

We build a point pattern made of cases (the points of interest) and controls (the background distribution of points).

Cases are a Matérn [@Matern1960] point pattern with $\kappa$ (expected) clusters of $\mu$ (expected) points in a circle of radius *scale*.
Controls are a Poisson point pattern whose density $\lambda$ decreases exponentially along the y-axis (we will call "north" the higher y values).

```{r}
library(dbmss)
# Simulation of cases (clusters)
rMatClust(kappa = 10, scale = 0.05, mu = 10) %>% 
  as.wmppp -> 
  CASES
CASES$marks$PointType <- "Cases"
# Number of points
CASES$n

# Simulation of controls (random distribution)
rpoispp(function(x, y) {1000 * exp(-2 * y)}) %>% 
  as.wmppp ->
  CONTROLS
CONTROLS$marks$PointType <-"Controls"
# Number of points
CONTROLS$n

# Mixed patterns (cases and controls)
ALL <- superimpose(CASES, CONTROLS)
autoplot(ALL)
```


## Calculate and plot *M* Cases

```{r}
# Fix the number of simulations and the level of risk
NumberOfSimulations <- 100
Alpha <- .01

# Calculate and plot M Cases
ALL %>% 
  MEnvelope(
    ReferenceType="Cases", 
    SimulationType = "RandomLocation",
    NumberOfSimulations = NumberOfSimulations,
    Alpha = Alpha,
    Global = TRUE
  ) ->
  M_env_cases

autoplot(M_env_cases)
```

The plot shows a clear relative concentration of cases.


## Map *M* results

To plot the individual values of *M* around each case, a distance must be chosen.
Then, the function must be computed at this distance with individual values.
Finally, a kriged weighted, marked, planar point patterns (`kwmppp`) object is produced and plotted.

```{r}
# Choose the distance to plot
Distance <- 0.1

# Calculate the M values to plot
ALL %>% 
  Mhat(
    r = c(0, Distance),
    ReferenceType = "Cases",
    NeighborType = "Cases", 
    Individual = TRUE
  ) -> 
  M_TheoEx

# Map resolution
resolution <- 512

# Create a map by smoothing the local values of M
M_TheoEx_map <- Smooth(
  ALL, 
  fvind = M_TheoEx, 
  distance = Distance,
  ReferenceType ="Cases",
  Nbx = resolution, Nby = resolution
)

# Plot the point pattern with values of M(Distance)
plot(M_TheoEx_map)
# Add the cases to the map
points(
  ALL[ALL$marks$PointType == "Cases"],
  pch = 20, col = "red"
)
# Add contour lines
contour(M_TheoEx_map, add=TRUE)
```

We can see that cases are concentrated almost everywhere (local *M* value above 1) because we chose a Matérn point pattern.

The areas with the higher relative concentration are located in the north of the map because the controls are less dense there.


## Compare with the density of cases

The density of cases is plotted.
High densities are *not* similar to high relative concentrations in this example because the control points are not homogeneously distributed.

```{r}
plot(density(CASES), main = "")
points(
  ALL[ALL$marks$PointType == "Cases"],
  pch = 20, col = "red"
)
```


## Confidence level

The individual values of *M* can be compared to their distribution under the null hypothesis of radom location, i.e. where all points except for the reference one are shuffled among all actual locations.

The quantiles of the actual values with respect to their null hypothesis distribution can is mapped.

```{r}
# Calculate the M values to plot with their confidence level
ALL %>% 
  Mhat(
    r = c(0, Distance),
    ReferenceType = "Cases",
    NeighborType = "Cases", 
    Individual = TRUE,
    Quantiles = TRUE,
    NumberOfSimulations = NumberOfSimulations,
    Alpha = Alpha
  ) -> 
  M_TheoEx

# Create a map of the quantiles
M_TheoEx_q_map <- Smooth(
  ALL, 
  fvind = M_TheoEx, 
  distance = Distance,
  ReferenceType ="Cases",
  Quantiles = TRUE,
  Nbx = resolution, Nby = resolution
)

# Plot the point pattern with values of M(Distance)
plot(M_TheoEx_q_map)
# Add the cases to the map
points(
  ALL[ALL$marks$PointType == "Cases"],
  pch = 20, col = "red"
)
# Add contour lines of the critical values
contour(M_TheoEx_map, levels = c(Alpha/2, 1 - Alpha/2), add=TRUE)
```

The contour lines show the regions where the local *M* value is smaller of greater than the critical values of the simulations, i.e. it is among the 1% extreme values.


# Suzanne Lenglen Park

## Data

Our data is extracted from "Paris open data" [^1].

[^1]: https://opendata.paris.fr.


### Data wrangling

Data are stored in `trees_2021.zip` which contains two GeoJSON files:

- trees_2021 stores all trees of the city of Paris in 2021.
- trees_logged contains all trees logged after 2021.

They must be read.
Data are projected into the Lambert 93 datum so that coordinates are in meters.

```{r data}
unzip("data/trees_2021.zip", exdir = "data")
library("sf")
read_sf("data/trees_2021.geojson") %>% 
  st_transform(crs = 2154) -> 
  trees_all_raw
read_sf("data/trees_logged.geojson") %>% 
  st_transform(crs = 2154) -> 
  trees_logged_raw
```


#### All trees

The first dataset contains all trees in Paris in 2021, including those to be cut.

Trees from the Suzanne Lenglen park are selected.
Columns of interest are:

- `ID`: a numeric unique identifier for each tree.
- `Species_name`: the scientific name of the tree species, i.e. Genus species.
- `Status`: Alive. 
- `Genus`.
- `Species`.
- `French_species_name`: vernacular name.
- `Circumference`: in cm.

```{r}
library("dplyr")
trees_all_raw %>% 
  # Filter Suzanne Lenglen park
  filter(adresse == "PARC OMNISPORT SUZANNE LENGLEN / 7 BOULEVARD DES FRERES VOISIN") %>% 
  # Create a field with the species name
  mutate(Species_name = as.factor(paste(genre, espece))) %>%
  # Create a field with the status
  mutate(Status = "Alive") %>%
  # Genus and Species fields
  mutate(Genus = as.factor(genre)) %>% 
  mutate(Species = as.factor(espece)) %>% 
  # Rename and finally select columns
  rename(
    ID = idbase,
    French_species_name = libellefrancais,
    Circumference = circonferenceencm
  ) %>% 
  select(
    ID, 
    Species_name, 
    Status, 
    Genus, 
    Species, 
    French_species_name,
    Circumference) -> 
  trees_all

# Number of trees
trees_all %>% nrow()
```

We have `r nrow(trees_all)` trees in the park.


#### Logged trees

Logged trees are in the second dataset.

Their status is "Logged".
An extra field, `Logging_reason` contains the motivation to cut them off (in French).
`Circumference` is absent.


```{r}
# Tree description
trees_logged_raw %>%
  # Filter Suzanne Lenglen park
  filter(adresse == "PARC OMNISPORT SUZANNE LENGLEN / 7 BOULEVARD DES FRERES VOISIN") %>% 
  # Exclude unidentified trees
  filter(!is.na(especearbreprecedent), !is.na(libellefrancaisarbreprecedent), !is.na(genrearbreprecedent)) %>%
  filter(libellefrancaisarbreprecedent != "Non spécifié") %>%
  filter(especearbreprecedent != "n. sp.") %>% 
  # Create a field with the species name
  mutate(Species_name = as.factor(paste(genrearbreprecedent, especearbreprecedent))) %>%
  # Create a numeric ID
  mutate(ID = as.integer(idbase)) %>%
  # Create a field with the status
  mutate(Status = "Logged") %>%
  # Genus and Species fields
  mutate(Genus = as.factor(genrearbreprecedent)) %>% 
  mutate(Species = as.factor(especearbreprecedent)) %>% 
  # Reason for logging (in French)
  mutate(Logging_reason = motifabattagearbreprecedent) %>% 
  # Rename and finally select columns
  rename(
    French_species_name = libellefrancaisarbreprecedent
  ) %>% 
  select(
    ID, 
    Species_name, 
    Status, 
    Genus, 
    Species, 
    Logging_reason, 
    French_species_name
  ) -> 
  trees_logged

# Number of trees
trees_logged %>% nrow()
```

`r nrow(trees_logged)` among the `r nrow(trees_all)` trees of the park were logged.


#### Merge

The two datasets are merged here.

The logged trees must be removed from the first one.
`Circumference` is removed because it is missing from the logged trees dataset.

```{r}
# All trees
trees_all %>%
  # Delete the logged trees
  filter(!(ID %in% trees_logged$ID)) %>% 
  # Delete the circumference that is absent in trees_logged
  mutate(Circumference = NULL) %>% 
  # Bind the logged trees
  bind_rows(trees_logged) -> 
  trees_no_circumference
```

Circumferences of all trees, including logged ones, are in `tree_all` from where they can be recovered.

```{r}
# Prepare a tibble with circumferences
trees_all %>%
  select(ID, Circumference) %>% 
  # inner_join.sf refuses sf objects
  st_set_geometry(NULL) ->
  Circumferences
# Add the Circumference of trees
trees_no_circumference %>%
  inner_join(Circumferences, by = "ID") ->
  trees
```


#### Shorter logging reasons

Logging reasons can be:

- Decaying: the tree's condition is not healthy enough to keep it safely in a public park.
- infested: the tree is a maple affected by the (contagious) sooty bark disease, caused by the fungus *Cryptostroma corticale*.

```{r}
library("stringr")
trees$Logging_reason[is.na(trees$Logging_reason)] <- ""
trees$Logging_reason %>% 
  str_replace("Arbre.*", "Decaying") %>% 
  str_replace("Foyer.*", "infested") ->
  trees$Logging_reason
```


#### Factors

Several fields are converted to factors for efficiency.

```{r}
trees$Logging_reason <- as.factor(trees$Logging_reason)
trees$Status <- as.factor(trees$Status)
trees$French_species_name <- as.factor(trees$French_species_name)
```


### Point patterns

dbmms uses weighted, marked, planar point patterns (wmppp).
A wmppp named `trees_infested` is built. Point marks are their basal area (as weight) and either their logging reason or their genus if they are alive. 


```{r}
library("dbmss")
trees %>%
  # Weight is the basal area
  mutate(PointWeight = Circumference^2 / 4 / pi) %>%
  mutate(
    PointType = ifelse(
      Logging_reason == "", 
      as.character(Genus), 
      as.character(Logging_reason)
    )
  ) %>%
  # Add X and Y
  bind_cols(st_coordinates(trees)) %>% 
  wmppp(
    window = as.owin(st_bbox(trees)),
    unitname = c("meter", "meters")
    ) ->
  trees_infested
```

We also need a point pattern to describe the park before logging, as a reference.

```{r}
trees_all %>%
  # Weight is the basal area
  mutate(PointWeight = Circumference^2 / 4 / pi) %>%
  # Genus is the point type
  rename(PointType = Genus) %>%
  # Add X and Y
  bind_cols(st_coordinates(trees_all)) %>% 
  wmppp(
    window = as.owin(st_bbox(trees_all)),
    unitname = c("meter", "meters")
    ) ->
  trees_2021
autoplot(trees_2021) + 
  ggplot2::labs(size="Basal area", color="Genus")
```

The map shows the tree genera. 
Maple (*Acer spp.*) are the most abundant trees in the park.


## Spatial concentration of maple trees

The *M* statistic is computed to detect the spatial concentration of maple trees before logging. 

```{r}
Distance <- 15
trees_2021 %>% 
   MEnvelope(
     r = 0:(10 * Distance) / 5,
     ReferenceType = "Acer",
     NeighborType = "Acer", 
     NumberOfSimulations = NumberOfSimulations
  ) ->
  M_Acer

autoplot(M_Acer)
```

To map it, individual values must be calculated at the chosen distance, that is 15 meters.

```{r}
trees_2021 %>% 
   Mhat(
     r = c(0, Distance),
     ReferenceType = "Acer",
     NeighborType = "Acer", 
     Individual = TRUE
  ) ->
  M_ind_Acer
```

### Smoothed map

A map is produced by smoothing the individual values.
To respect the geometry of the map, the ratio of rows to columns of the image to produced is first calculated.

```{r}
# Window ratio
ratio <- with(
  trees_infested$window, 
  {
   (yrange[2] - yrange[1]) / (xrange[2] - xrange[1])
  }
)

# Smooth the values
trees_2021 %>% 
  Smooth(
    fvind = M_ind_Acer, 
    distance = Distance, 
    ReferenceType = "Acer",
    Quantiles = FALSE,
    Adjust = 4,
    Nbx = resolution, Nby = resolution * ratio
  ) -> map_acer

plot(map_acer)
# Add contour lines
contour(map_acer, nlevels = 10, add=TRUE)
# Add infested trees
points(
  trees_infested[trees_infested$marks$PointType == "infested"],
  pch = 20
)
# And decaying trees
points(
  trees_infested[trees_infested$marks$PointType == "Decaying"],
  pch = 4,
  col = "red"
)
```


### Kriged map

The map requires krigeing the individual values on a grid of points.
To build the grid, the size ratio of the spatial window is calculated.
The number of rows and columns of the grid will respect this ratio so that its points are equally spaced.

```{r}
# Window ratio
ratio <- with(
  trees_infested$window, 
  {
   (yrange[2] - yrange[1]) / (xrange[2] - xrange[1])
  }
)
# Map resolution: number of columns of the grid.
resolution <- 512
```

A kriged weighted, marked, planar point patterns (`kwmppp`) object is produced and plotted.
Logged trees are added to the map:

- infested trees are black points,
- Decaying trees are red crosses.

```{r}
trees_2021 %>% 
  kwmppp(
    fvind = M_ind_Acer,
    distance = Distance,
    ReferenceType = "Acer",
    Nbx = resolution,
    Nby = resolution * ratio
  ) ->
map_acer

plot(map_acer)
# Add infested trees
points(
  trees_infested[trees_infested$marks$PointType == "infested"],
  pch = 20
)
# And decaying trees
points(
  trees_infested[trees_infested$marks$PointType == "Decaying"],
  pch = 4,
  col = "red"
)
```

infested trees are present in the areas where maples are concentrated.

The 15-meter distance has been chosen because it is a plausible distance of contagion between trees.
A shorter distance, say 6 meters, would be justified as a peak of *M*(distance) but neighbors are scarse at this scale, leading to less statistical power.
At greater distances, such as 30m, concentration is still significant but neighborhoods overlap more, leading to less clear results.

Both alternative distances are tested here.

Considering 6-meter-radius neighborhoods, the map shows that some high and low values of local concentration are not detected:

```{r, echo = FALSE}
trees_2021 %>% 
   Mhat(
     r = c(0, 6),
     ReferenceType = "Acer",
     NeighborType = "Acer", 
     Individual = TRUE
  ) ->
  M_ind_Acer_6
trees_2021 %>% 
  kwmppp(
    fvind = M_ind_Acer_6,
    distance = 6,
    ReferenceType = "Acer",
    Nbx = resolution,
    Nby = resolution * ratio
  ) ->
map_acer_6

plot(map_acer_6)
# Add infested trees
points(
  trees_infested[trees_infested$marks$PointType == "infested"],
  pch = 20
)
# And decaying trees
points(
  trees_infested[trees_infested$marks$PointType == "Decaying"],
  pch = 4,
  col = "red"
)
```

30-meter-radius neighborhoods are shown below:

```{r, echo = FALSE}
trees_2021 %>% 
   Mhat(
     r = c(0, 30),
     ReferenceType = "Acer",
     NeighborType = "Acer", 
     Individual = TRUE
  ) ->
  M_ind_Acer_30
trees_2021 %>% 
  kwmppp(
    fvind = M_ind_Acer_30,
    distance = 30,
    ReferenceType = "Acer",
    Nbx = resolution,
    Nby = resolution * ratio
  ) ->
map_acer_30

plot(map_acer_30)
# Add infested trees
points(
  trees_infested[trees_infested$marks$PointType == "infested"],
  pch = 20
)
# And decaying trees
points(
  trees_infested[trees_infested$marks$PointType == "Decaying"],
  pch = 4,
  col = "red"
)
```

Increasing the distance that defines neighborhoods extends the areas where high or low concentration is detected around the most isolated trees, making them more visible, but blurs them in dense areas, where high and low values overlap and cancel out (e.g. in the south-eastern part of the map).
So the appropriate distance should be chosen according to the knowledge of the studied process.


## Concentration of maples around infested trees

To test the intertype concentration between sane and infested maple trees, the following $M$-intertype is computed: we evaluate whether the proportion of maples trees is more important in the close neighborhood of infested maples than elsewhere. 
If so, results give support to the existence of a contagious disease [@Hantsch2014].

```{r}
trees_infested %>% 
   MEnvelope(
     r = 0:(10 * Distance) / 5,
     ReferenceType = "infested",
     NeighborType = "Acer",
     NumberOfSimulations = NumberOfSimulations
  ) ->
  M_infested_Acer

autoplot(M_infested_Acer)
```

The *M* plot detects a significant concentration of maples around infested maples.

The map is produced.

```{r}
trees_infested %>% 
   Mhat(
     ReferenceType = "infested",
     NeighborType = "Acer", 
     Individual = TRUE
  ) ->
  M_ind_infested_Acer

trees_infested %>% 
  kwmppp(
    fvind = M_ind_infested_Acer,
    distance = Distance,
    ReferenceType = "infested",
    Nbx = resolution,
    Nby = resolution * ratio
  ) ->
map_infested_acer

plot(map_infested_acer)
# Add infested trees
points(
  trees_infested[trees_infested$marks$PointType == "infested"],
  pch = 20
)
# And decaying trees
points(
  trees_infested[trees_infested$marks$PointType == "Decaying"],
  pch = 4,
  col = "red"
)
```

The map shows that the local concentration of maple around infested trees is generally high (warm colors around black points).

`r if (!knitr:::is_latex_output()) '# References {-}'`
